---
title: 'Clustering our primary and CL samples and replotting heatmap. '
output:
  html_document:
    df_print: paged
---

I will use the data where Jaime has already removed the batch effects. The script he has pointed me to is: `/shared/sudlab1/General/projects/PC_MM_Atac_seq/ians/Code/ATAC_clustering_MM_subgroup_DA_reg_enh_our_primary_rlog_batch_corr_acc_subg_with_jin_et_al_samples_rlog_batch_corr_not_acc_subg_12_08_2020.R`. 

Start my reading in the rlogs

```{r}
# Data originally in /fastdata/mbp15ja/jin_et_al_primary_samples_ATAC_clustering_MM_subgroup_vs_PC_cons_peaks_07_08_2020/rlog_batch_correct_read_counts_in_cons_peaks_jin_et_al_2018_and_my_study_jin_et_al_not_acc_condition.tsv.gz
# Copied to /shared/sudlab1/General/projects/PC_MM_Atac_seq/ians/data/rlog_batch_correct_read_counts_in_cons_peaks_jin_et_al_2018_and_my_study_jin_et_al_not_acc_condition.tsv.gz
corrected_rlog_data <- read.delim("../data/rlog_batch_correct_read_counts_in_cons_peaks_jin_et_al_2018_and_my_study_jin_et_al_not_acc_condition.tsv.gz")

```

Get the regions we want to consider:

```{r}
library(dplyr)
regions <- read.delim("../data/new_subgroup_MM_vs_PC_DE_ATAC_DE_RNA_1Mb.tsv.gz", stringsAsFactors = FALSE) %>%
  mutate(region_id = paste0(chr,":",start,"-",end))

```

Now subset the rlogs so that it only has the signifciant regions of interest

```{r}
sig_rlogs <- regions %>%
  select(region_id) %>%
  inner_join(corrected_rlog_data, by = c("region_id" = "id"))

sig_rlogs <- corrected_rlog_data %>%
  filter(id %in% regions$region_id) %>%
  arrange(id)
```

The next step is the column data, which is stored in a file in the temporary directory. 

```{r}

col_data <- read.delim("../data/coldata_ATAC_yin_our_primary.tsv.gz",
                       stringsAsFactors = FALSE)  %>%
  mutate(X=ifelse(grepl("^[0-9]", X), paste0("X",X), X)) %>% 
  mutate(X= gsub(",", ".", X, fixed=TRUE)) %>%
  mutate(study = ifelse(batch == "batch_1", "This study", "Jin et al"))

```

Now I need to cluster the columns of the rlog matrix


```{r}

mat <- as.matrix(sig_rlogs[,-1])
col_hust <- hclust(as.dist(1-cor(mat)), method="ward.D2")

row_hclust <- hclust(as.dist(1-cor(t(mat))), method="ward.D2")
```

Now plot the Rlog heatmap. 

```{r}
library(ComplexHeatmap)
library(circlize)
cbbPalette =  c("ND"="#000000", "MAF"="#E69F00", "CCND1"="#56B4E9", "MMSET"="#009E73", "#F0E442", 
                "CD19- NPC"="grey40", "HD"="#0072B2", "#D55E00", "UNKNOWN"="#CC79A7", "0"="white")
study_pal = c("This study" = "#7ac57b", "Jin et al" = "#58674f")
col_fun = colorRamp2(c(-3.5, 0, 3.5), c("blue", "white", "red"))
column_ha = HeatmapAnnotation(Subtype=col_data$condition,
                              Study=col_data$study,
                              which="column",
                              col=list(Subtype=cbbPalette,
                                       Study=study_pal), annotation_name_gp = gpar(fontsize=8), 
                              show_legend=c(FALSE, TRUE), 
                              show_annotation_name = FALSE,
                              annotation_height = c(unit(0.1, "cm"), unit(0.1,"cm")),
                              annotation_legend_param = list(labels_gp=gpar(fontsize=8), title_gp=gpar(fontsize=8)))
png("../Figures/ATAC_jin_et_al_heatmap_no_LFC.png", width=6, height=4, units = "in",  res=600)
 Heatmap(t(scale(t(mat))),
  #                      col = col_fun,
        name="ATAC\nsignal",
        cluster_rows = row_hclust,
        cluster_columns = col_hust,
        row_dend_reorder=FALSE,
        column_dend_reorder=FALSE,
        show_row_dend = FALSE,
        show_column_names = FALSE,
        top_annotation = column_ha,
        heatmap_legend_param= list(labels_gp = gpar(fontsize=8), title_gp=gpar(fontsize=8)))

dev.off()
```

Finally, Figure 3A had the LFC matrix, and the rows of the heatmap was ordered by them. 


```{r}
sig_matrix <- regions %>%
  select_at(vars(contains("ATAC_DE"), contains("region_id"))) %>%
  distinct() %>%
  arrange(region_id)
# sig_matrix = regions[,grepl("ATAC_DE", colnames(regions))]
colnames(sig_matrix) <- sub("ATAC_DE_", "", colnames(sig_matrix))
colnames(sig_matrix) <- sub("_vs_ND", "", colnames(sig_matrix))
#colnames(sig_matrix) <- sub("HD", "HY", colnames(sig_matrix))
sig_matrix <- sig_matrix[,c("MAF", "MMSET", "CCND1", "HD")]
for (subtype in colnames(sig_matrix)) {
  sig_matrix[,subtype][sig_matrix[,subtype] != 1] <- "0"
  sig_matrix[,subtype][sig_matrix[,subtype] == 1] <- subtype

}
row_hclust <- hclust(dist(sig_matrix != "0"), method="average")
sig_heat <- Heatmap(as.matrix(sig_matrix),
                    col=cbbPalette,
                    cluster_columns=FALSE,
                    show_heatmap_legend = FALSE,
                    column_names_side = "top",
                    column_names_gp = gpar(fontsize=8))
sig_heat
```

```{r, fig.height=3, fig.width=5}
regions_heat <- Heatmap(t(scale(t(mat))),
  #                      col = col_fun,
        name="ATAC\nsignal",
        cluster_rows = row_hclust,
        cluster_columns = col_hust,
        row_dend_reorder=FALSE,
        column_dend_reorder=FALSE,
        show_row_dend = FALSE,
        show_column_names = FALSE,
        top_annotation = column_ha,
        heatmap_legend_param= list(labels_gp = gpar(fontsize=8), title_gp=gpar(fontsize=8)))

regions_heat + sig_heat # -> combined_heatmap

png("../Figures/jin_et_heatmap.png", height=3, width=5, units="in", res = 300)
regions_heat + sig_heat 
dev.off()

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
